---
title: "Homework: Reading in Data"
author: "Firstname Lastname"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

saspath <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en" # Your SAS path goes here
sasopts <- "-ls 75"
knitr::opts_chunk$set(engine.path=list(sas=saspath, saslog=saspath, sashtml=saspath),
                      engine.opts=list(sas=sasopts, saslog=sasopts, sashtml=sasopts), 
                      comment=NA)
library(SASmarkdown)
```

## Instructions
Big picture: Read in the Medicare and Medicaid General Payments Data and conduct an exploratory data analysis. You should use both SAS and R to explore the data, but you might choose to use R for certain tasks and SAS for other tasks. 

When you are finished with the assignment: 

1. Save the file as 04_Lastname_Firstname.Rmd and compile it
2. Commit the Rmd file and corresponding html file to your homework git repo
3. Push the commit to github
4. Locate the Rmd file on github and click on the SHA link for the file
5. Paste the url of the resulting page into Canvas to submit your assignment.

Your assignment must compile on a different computer as long as the saspath is set correctly for that machine. This means you will need to use a local file path when you read the data in via R. 

### About the Data

To get this data, I started with the full 2018 dataset, and limited it to cash payments made to medical doctors reported in January. The full dataset is *way* too big - several GB when stored on disk. 

A sample of the full dataset is available in the template repository (and also  [here](https://github.com/srvanderplas/unl-stat850/raw/master/data/General_Payment_Data_Sample.csv)). It's 8MB after extreme trimming of the less useful columns and sampling 25% of the rows. You can read about the data [here](https://openpaymentsdata.cms.gov/about) and [here](https://www.cms.gov/openpayments/). 

### Questions to Address

- Describe the dataset and any interesting things you find within. 
- What tasks are easier in R? SAS?
- What do you find that might need to be cleaned up or corrected before analysis?

You may want to include graphics (using the sample code in the book, or the [R Graph Gallery](https://www.r-graph-gallery.com/)) to show any interesting things you discover. When you include a graph, be sure to provide some contextual description of the information you want someone to take away from the graph.

## Data Exploration

### R
(You don't need to keep these headings, but I want you to have a skeleton of what the code chunks should look like)

```{r}
data <- read_csv("https://github.com/srvanderplas/unl-stat850/raw/master/data/General_Payment_Data_Sample.csv")

```

### SAS
Fill in your code in the chunk below.
```{sashtml data-import, error = T, collectcode=T}
filename dat "General_Payment_Data_Sample.csv";

PROC IMPORT DATAFILE = dat OUT = openmedpay 
    DBMS = CSV 
    REPLACE; 
    GETNAMES = YES;
    GUESSINGROWS = 3500;
RUN;

```

You can delete the information below if you don't need it, but these are handy tricks for using sasmarkdown. 
```{sashtml, collectcode=T}
/* The collectcode=T option lets you reference data in other chunks, just like in R */
  
```


```{sas, eval = F, collectcode=T}
/* If you want to print the SAS log out, use these chunks */
/* Note that this is in a sas chunk, but the R chunk below actually prints */
/* the log file in Rmarkdown */
  
proc printto log="saslog.log" new; /* This comes at the start of the chunk */

/* You can use this sequence of chunks as many times as you want in a document */
```

```{r}
cat(readLines("saslog.log"), sep="\n")
# Do not forget to remove the log file when you are done!
unlink("saslog.log")
```

